A Quantitative Genetic Linked Selection Model of the Negative Selection Process in Humans

### 2023-07-11 manuscript/manuscript.tex
working in andys edits

\begin{abstract}

    Previous work has shown that negative selection against the steady influx
    of deleterious mutations is the predominant process shaping genome-wide
    patterns of genetic diversity in humans. This process leaves a ``linked
    selection signal"

    The classic population genetic
    model of this process, background selection, considers

    % TODO: emphasize connection to nearly neutral theory

    % uniting quant gen/fitness variation, linked selection, and substitution
    % The interesting thing about connecting all these models is that we might
    % have hope of calibrating these models to the true process.
\end{abstract}

@@

The continual influx of new mutations into populations is the ultimate source
of all adaptations, but the vast majority of mutations either do not affect
fitness or are deleterious. Negative selection (i.e. purifying selection) works
to eliminate such deleterious mutations from the population, which is essential
to maintaining the integrity of information in the genome needed to maintain a
lineage. Mutations under negative selection are kept at low frequencies
(Haldane?), contributing to human traits (XXX) and diseases. Moreover, the
negative selection process determines sequence conservation across phylogenetic
timescales, which is evident in genome-wide cross-species comparisons
\parencite{Siepel2005-wh}. As with most selection in the genome, the negative
selection process stochastically perturbs the allele frequencies of neighboring
linked variation. This reduces genetic variability to create a large-scale
spatial linked selection ``reduction signal" in genome-wide diversity. Mutation
and negative selection's contribution to this signal is mediated by the
deleterious mutation rate, the strength of selection in conserved segments, and
the spatial distribution of recombination rates and conserved segments. Since
genome-wide recombination maps and putatively conserved segments are available
for many species, researchers have combined these with population genetic
theory to estimate the mutation rate and selection coefficients of deleterious
mutations, as well as the overall reduction in genetic variability due to the
negative selection process.

>>>>

\begin{abstract}
    Across the human genome, there are large-scale fluctuations in genetic
    diversity caused by the indirect effects of selection. This can be thought
    of as a ``linked selection signal" that reflects selection's effects
    varying according to the placement of functional regions and recombination
    rates along the genome. Previous work has shown that negative selection
    against the steady influx of new deleterious mutations into conserved
    regions is the predominant mode of selection in humans. However, the
    theoretic model that underpins these results, classic Background Selection
    theory, is only applicable when new mutations are so deleterious that they
    cannot drift up in frequency and ultimately fix in the population. Here, we
    develop an alternate theoretic basis for genome-wide linked selection based
    on a quantitative genetics view, which considers the distribution of
    additive fitness variance along the genome. Using a recent model that
    jointly predicts the equilibrium fitness variance and substitution rates
    due to deleterious mutations, we estimate the distribution of selection
    coefficients and mutation rate across three human populations. While our
    model can accommodate weaker selection, we find consistent evidence across
    all populations of very strong selection against deleterious mutations. We
    also find that our model's predicted substitution rates are qualitatively
    consistent with observed divergences in features under weak selective
    constraint like introns and UTRs, but less so in regions under strong
    selective constraint. While our model explains up to 67\% of the out-sample
    variance in diversity at the megabase scale, we find evidence of residual
    strong selection against loss-of-function variants unfit by our model,
    indicating the dominant role negative selection plays in the human genome.
    Altogether, our genome-wide model of the linked selection is a step towards
    uniting evolutionary quantitative genetics with substitution processes.
\end{abstract}

@@

The continual influx of new mutations into populations is the ultimate source
of all adaptations, but the vast majority of mutations either do not affect
fitness or are deleterious. The best predictor of whether a mutation will be
harmful to fitness is if it lands in a functional region, which are often
conserved over deep phylogenetic timescales \parencite{Siepel2005-wh}. Since
such high conservation reflects the product of millions or even billions of
years of evolutionary optimization, it follows that deleterious variants should
be the overwhelming majority of segregating variation in functional regions,
and be responsible for a significant portion of phenotypic variation and
disease \parencite{Zeng2018-ci} (XXX).

Both positive and negative selection stochastically perturb the allele
frequencies of neighboring linked variation in a phenomenon known as linked
selection \parencite{Nordborg1996-nq,Maynard_Smith1974-zr,Barton1998-aj}. While
linked selection processes like sweeps can leave signals on small spatial
scales (XXX), the precise arrangement of functional regions and recombination
rates across the genome leads to a large-scale linked selection signal.

Mutation and negative selection's contribution to this signal is mediated by
the deleterious mutation rate, the strength of selection in conserved segments,
and the spatial distribution of recombination rates and conserved segments.
Since genome-wide recombination maps and putatively conserved segments are
available for many species, researchers have combined these with population
genetic theory to estimate the mutation rate and selection coefficients of
deleterious mutations, as well as the overall reduction in genetic variability
due to the negative selection process.

HERE

As deleterious variation is clustered in functional portions of the genome,
we expect linked selection to reduce levels of diversity surrounding evolutionarily
conserved segments (e.g. genes, regulatory elements, etc.) 
and to shape large-scale spatial patterns in genetic diversity. 
This landscape of diversity is the product of a complex interplay between the
deleterious mutation rate, the local recombination rate, the strength of selection, and
the spatial distribution of conserved segments along a chromosome. 
Since genome-wide recombination maps and putatively conserved segments are available
for many species, there has been consistent effort to fit models of
linked selection to patterns of diversity to estimate population genetic
parameters such as the strength of selection and the deleterious mutation rate 
\parencite{Hudson1995-pt,McVicker2009-ax},
the overall reduction in genetic variability due to the
selection on deleterious variants \parencite{Elyashiv2016-vt,Murphy2022-sj},
and in some cases the rate of beneficial mutations \parencite{corbett2015natural,Elyashiv2016-vt}.
